### 2015-06-12 src/4_OnlineLearning.tex
pequeños detalles
demo variance and competitive analysis

\begin{equation*}
	\mathbf{C}=
	\begin{bmatrix} \quad &\mathbf{A}& \quad \\ \hdashline \quad
& \lambda\mathbb{I}&\quad  
\end{bmatrix} 
\quad \text{and} \quad
	\mathbf{F}=\begin{bmatrix} \quad &\mathbf{Y}& \quad \\ \hdashline \quad
& \mathbf{0}&\quad  \end{bmatrix} \, .
\end{equation*}

@@

\begin{eqnarray*}
\mathbf{X}(\lambda) &=&( \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1}\mathbf{Y} \\
&=& (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1} (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{Y} \\
&=&  (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1}  \hat{\mathbf{X}} \\
&=& \mathbf{W} \hat{\mathbf{X}} 
\end{eqnarray*}

@@

\begin{eqnarray*}
Var(\mathbf{X}(\lambda)) &=& Var(\mathbf{W}\hat{\mathbf{X}}) \\ 
&=& \mathbf{W}Var(\hat{\mathbf{X}})\mathbf{W}^\top \\
&=& \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top
\end{eqnarray*}

The figure~\ref{fig:biasvariance} shows the bias-variance tradeoff
given by equation~(\ref{eq:prederror}). Ridge regression shows an
increasing bias and a decreasing variance. Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}.

@@

In principle, the stream of observations may be infinitely long, or the next
observation long delayed, precluding any hope of waiting until all the
observations have been received. One would rather not simply accumulate and
store all the inputs and, upon receipt of each new one, apply a batch learning
algorithm to the entire sequence of inputs received so far. It would be
preferable compu- tationally if the existing hypothesis or other artifact of
learning could be updated in response to each newly received input observation.

Incremental learning is very useful when there is no need to record fundamental
data and only a summary needs to be retained. Due to this, incremental
algorithms are often characterized as memoryless, because no memory of past data
is required.  The algorithm is online but not incremental if it doesn't produce
the same result for all observations that the corresponding batch algorithm
would for these same observations.

@@

@@

The AAR, proposed by~\cite{vovk2001}, is an application of the aggregating
algorithm to the problem of regression. The idea is introduce the new input
vector $\mathbf{x}_{m+1}$ to solve the model parameters:

>>>>

\begin{equation*}
	\mathbf{C}=
	\begin{bmatrix} \quad &\mathbf{A}& \quad \\ \hdashline \quad
& \sqrt\lambda\mathbb{I}&\quad  
\end{bmatrix} 
\quad \text{and} \quad
	\mathbf{F}=\begin{bmatrix} \quad &\mathbf{Y}& \quad \\ \hdashline \quad
& \mathbf{0}&\quad  \end{bmatrix} \, .
\end{equation*}

@@

\begin{eqnarray*}
\mathbf{X}(\lambda) &=&( \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1}\mathbf{A}^\top \mathbf{Y} \\
&=& (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1} (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{Y} \\
&=&  (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1}  \hat{\mathbf{X}} \\
&=& \mathbf{W} \hat{\mathbf{X}} 
\end{eqnarray*}

@@

\begin{eqnarray*}
Var(\mathbf{X}(\lambda)) &=& Var(\mathbf{W}\hat{\mathbf{X}}) \\
&=& E[(\mathbf{W}\hat{\mathbf{X}}-E[\mathbf{W}\hat{\mathbf{X}}])(\mathbf{W}\hat{\mathbf{X}}-E[\mathbf{W}\hat{\mathbf{X}}])^\top] \\
&=& \mathbf{W}E[(\hat{\mathbf{X}}-E[\hat{\mathbf{X}}])(\hat{\mathbf{X}}-E[\hat{\mathbf{X}}])^\top] \mathbf{W}^\top \\
&=& \mathbf{W}Var(\hat{\mathbf{X}})\mathbf{W}^\top \\
&=& \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top
\end{eqnarray*}

The figure~\ref{fig:biasvariance} shows the bias-variance tradeoff
given by equation~(\ref{eq:prederror}). Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}. Ridge regression shows an
increasing squared bias and a decreasing variance.

Since $Bias(\mathbf{X}(\lambda)) \neq 0$ this imply that

\begin{equation*}
Bias(\mathbf{X}(\lambda)) ^2 > 0 
\end{equation*}

\noindent we know that $\mathbf{A}^\top
\mathbf{A}$ has an eigenvalue decomposition $\mathbf{A}^\top
\mathbf{A} = \mathbf{V} \Sigma \mathbf{V}^{-1}$.

\begin{eqnarray*}
Var(\mathbf{X}(\lambda)) &=& \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top\\
Var(\mathbf{X}(\lambda)) &=& \sigma^2 (\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1} (\mathbf{A}^\top \mathbf{A} )^{-1}((\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1} )^\top \\
&=& \sigma^2 (\mathbb{I} + \lambda (\mathbf{V} \Sigma \mathbf{V}^{-1})^{-1} (\mathbf{V} \Sigma \mathbf{V}^{-1})^{-1}((\mathbb{I} + \lambda (\mathbf{V} \Sigma \mathbf{V}^{-1})^{-1})^{-1} )^\top \\
&=& \sigma^2 \mathbf{V} (\mathbb{I} + \lambda \Sigma^{-1})^{-1} \Sigma^{-1} (\mathbb{I} + \lambda \Sigma^{-1})^{-1}  \mathbf{V}^{-1}\\
&=& \sigma^2 \mathbf{V} \mathbf{D}^{-1} \mathbf{V}^{-1}
\end{eqnarray*}

\noindent where $\mathbf{D}^{-1}  = (\mathbb{I} + \lambda \Sigma^{-1})^{-1} \Sigma^{-1} (\mathbb{I} + \lambda \Sigma^{-1})^{-1}$ is a diagonal matrix with coefficients:

\begin{equation*}
d_i = \frac{\sigma_i^{-1}}{(1+\lambda\sigma_i^{-1})^2}
\end{equation*}

@@

In principle, the stream of observations may be infinitely long, or the next
observation long delayed, precluding any hope of waiting until all the
observations have been received. One would rather not simply accumulate and
store all the inputs and, upon receipt of each new one, apply a batch learning
algorithm to the entire sequence of inputs received so far. It would be
preferable computationally if the existing hypothesis or other artefact of
learning could be updated in response to each newly received input observation.

Incremental learning is very useful when there is no need to record fundamental
data and only a summary needs to be retained. Due to this, incremental
algorithms are often characterised as memoryless, because no memory of past data
is required.  The algorithm is online but not incremental if it doesn't produce
the same result for all observations that the corresponding batch algorithm
would for these same observations.

@@

\subsection{Competitive analysis}

Competitive analysis was designed for analysing the performance of an online algorithm compared with its optimal offline algorithm. An optimal offline algorithm can view the sequences of requests in advance. The effectiveness of an online algorithm \cite{} may be measured by its competitive ratio which is the worst-case ratio of the online algorithm and the the optimal offline algorithm.

@@

The AAR, proposed by~\cite{vovk2001} (also known as Vovk-Azoury-Warmuth predictor ~\cite{azoury2001}), is an application of the aggregating
algorithm to the problem of regression. The idea is introduce the new input
vector $\mathbf{x}_{m+1}$ to solve the model parameters:
